---
title: "Introduction to NONCA"
author: "R Packager"
date: "2023-10-27"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to NONCA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

Welcome to `NONCA`, an R package designed for comprehensive pharmacokinetic non-compartmental analysis. This package provides a suite of functions to perform various calculations essential for pharmacokinetic studies, including Area Under the Curve (AUC), maximum concentration (Cmax), time to Cmax (Tmax), half-life, clearance, and more.

# Installation

You can install the development version of `NONCA` from GitHub:

```r
# install.packages("devtools")
# devtools::install_github("your_username/NONCA")
```

# Quick Start

Here's a quick example to get you started with `NONCA`.

First, let's load the package:

```{r setup}
library(NONCA)
```

Now, let's consider a simple dataset with time and concentration measurements.

```{r example_data}
time <- c(0, 0.25, 0.5, 1, 2, 4, 6, 8, 12, 24)
concentration <- c(0, 10, 25, 40, 35, 20, 10, 5, 2, 0.5)
```

We can calculate the AUC from time 0 to the last observed time:

```{r auc_example}
# Example of AUC_linear_log_linear
# Assuming a data frame or similar structure is expected by the actual function
# This is a placeholder; adjust based on actual function signature
# nonca_data <- data.frame(time = time, concentration = concentration)
# auc_val <- AUC_linear_log_linear(nonca_data$time, nonca_data$concentration)
# print(paste("AUC (0-t):", auc_val))
```

Please refer to the function documentation for detailed usage.

# Features

`NONCA` offers a wide range of functions for non-compartmental analysis, including:

*   **AUC Calculations**:
    *   `AUC_linear_log_linear()`: Area Under the Curve using linear-log linear trapezoidal rule.
    *   `AUC_linear_up_log_down()`: Area Under the Curve using linear up and log down trapezoidal rule.
    *   `AUC_0_inf()`: Area Under the Curve extrapolated to infinity.
*   **Key Parameters**:
    *   `Cmax_Tmax()`: Calculation of maximum concentration (Cmax) and time to Cmax (Tmax).
    *   `lambda_z()`: Estimation of terminal elimination rate constant (lambda_z).
    *   `half_life()`: Calculation of half-life.
    *   `clearance()`: Calculation of clearance.
    *   `volume_of_distribution()`: Calculation of volume of distribution.
    *   `MRT()`: Mean Residence Time.
    *   `accumulation_ratio()`: Accumulation ratio.
    *   `bioavailability()`: Bioavailability.

# Usage Examples

Detailed usage examples for each function are provided in their respective help files. You can access them using `?function_name` in your R console.

# Data Format Requirements

Functions in `NONCA` typically expect numerical vectors for `time` and `concentration`. Specific requirements for each function are detailed in their documentation.

# Citation

If you use `NONCA` in your research, please cite it as follows:

```
# To be filled in once the package is published.
```
